---
title: "Examples for all package functions"
author: "Bodo Burger"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: yes
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Examples for all package functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(intame)
library(ggplot2)
set.seed(4218)
theme_set(theme_light())
```

## `intame()`

### Basic example

```{r}
library(nnet)
set.seed(4219)
n = 200
x = runif(n, min = 0, max = 1)
x1 = runif(n, min=0, max=1) + .5*x
x2 = runif(n, min=0, max=1) + .5*x
y2 = function(x) -4 * cos(4*pi*x) * x + 4
y = 4*x1 + y2(x2) + rnorm(n, 0, .5)
df = data.frame(y, x1, x2)
nnet.mod = nnet(y ~ ., df, size = 20, linout = TRUE,
  decay = 0.0001, maxit = 1000, trace = FALSE)
AME.x1 = intame(nnet.mod, df, "x1")
AME.x1
plot(AME.x1)
AME.x2 = intame(nnet.mod, df, "x2", fe_grid_size = 20)
AME.x2
plot(AME.x2)
```

### Binary classification with spam dataset

```{r, warning=FALSE}
spam.data = mlr::getTaskData(mlr::spam.task)
spam.fit = suppressWarnings(glm(type ~ ., data = spam.data, family = binomial(link = "logit")))
spam.predict = function(object, newdata) predict(object, newdata, type = "response")

feat.select = c("charDollar", "hpl", "people") 
spam.fit$coefficients[feat.select]
ame::computeAME(spam.fit, spam.data, feat.select, predict.fun = spam.predict)

intame.charDollar = intame(spam.fit, spam.data, "charDollar",
  predict_fun = spam.predict, fe_method = "ALE")
summary(intame.charDollar)
plot(intame.charDollar)

plot(computeALE(spam.fit, spam.data, "charDollar", predict_fun = spam.predict))
plot(computePD(spam.fit, spam.data, "charDollar", predict_fun = spam.predict))
plot(computePD(spam.fit, spam.data, "hpl", grid_size = 30,
  predict_fun = spam.predict))
plot(computePD(spam.fit, spam.data, "people", grid_size = 30,
  predict_fun = spam.predict))
```

### Quantile-based Average Marginal Effects

`intame()` allows you to calculate Average Marginal Effects for each decile 
(or another quantile) using `output_method = "AME"` and `x_splits = quantiles`.

```{r}
x_splits = quantile(df[, "x2"], probs = seq(0, 1, .1))[2:10] # do not use min and max
AME.x2.quantiles = intame(nnet.mod, df, "x2", fe_grid_size = 20, output_method = "AME",
  x_splits = x_splits)
AME.x2.quantiles
plot(AME.x2.quantiles)
```

## `computePD()`

### Regression example using nnet

```{r computePD-nnet}
require(nnet)
require(gridExtra)
n = 1000
x1 = runif(n, 0, 1)
x2 = runif(n, 0, 1)
x3 = runif(n, 0, 1)
y = x1 + 2 * x2^2 + (x1-0.5) * (x3-0.5) + rnorm(n, 0, 0.1)
df = data.frame(y, x1, x2, x3)
nnet.fit = nnet(y ~ ., data = df, size = 10, linout = TRUE,
  decay=0.01, maxit = 1000, trace = FALSE)
p1 = plot(computePD(nnet.fit, df, feature="x1", grid_size=50))
p2 = plot(computePD(nnet.fit, df, feature="x2", grid_size=50))
p3 = plot(computePD(nnet.fit, df, feature="x3", grid_size=50))
grid.arrange(p1, p2, p3, ncol=2)
```

## `computeALE()`

### Regression example using nnet

```{r}
require(nnet)
require(ggplot2)
require(gridExtra)
n = 1000
x1 = runif(n, 0, 1)
x2 = runif(n, 0, 1)
x3 = runif(n, 0, 1)
y = x1 + 2 * x2^2 + (x1-0.5) * (x3-0.5) + rnorm(n, 0, 0.1)
df = data.frame(y, x1, x2, x3)
nnet.fit = nnet(y ~ ., data = df, size = 10, linout = TRUE,
  decay=0.01, maxit = 1000, trace = FALSE)
predict_fun = function(object, newdata)
  as.numeric(predict(object, newdata, type="raw"))
p1 = plot(computeALE(nnet.fit, df, feature="x1", grid_size=50,
  predict_fun = predict_fun))
p2 = plot(computeALE(nnet.fit, df, feature="x2", grid_size=50,
  predict_fun = predict_fun))
p3 = plot(computeALE(nnet.fit, df, feature="x3", grid_size=50,
  predict_fun = predict_fun))
grid.arrange(p1, p2, p3, ncol=2)
```

## `partition()`

### Toy Example

```{r}
set.seed(0815)
n = 6
x = sample(n) + 1.5
y = runif(n, 0, 10)
df = data.frame(x, y)
breaks.rpart = partition(df$x, df$y, intervals = 4)
breaks.cluster = partition(df$x, df$y, intervals = 4,
  part.method = "cluster")
ggplot() + 
  geom_point(data=df, aes(x = x, y = y)) +
  geom_vline(aes(xintercept = breaks.rpart-.03,
    color = "rpart")) +
  geom_vline(aes(xintercept = breaks.cluster+.03,
    color = "cluster"))
```

### More complex example

```{r}
set.seed(123)
n = 100
x = runif(n, 0, 1000)
y = x * sin(.01*x) + rnorm(n, 0, 10)
df = data.frame(x, y)
poly12.mod = lm(y ~ poly(x, 12, raw = TRUE), data = df)
me = derivative(x, "x", df, poly12.mod)
breaks.rpart = partition(df$x, me, 5)
breaks.cluster = partition(df$x, me, 5, part.method = "cluster")
ggplot() + geom_line(data=df, mapping=aes(x = x, y = y)) +
  geom_vline(aes(xintercept = breaks.rpart+3, color = "rpart")) +
  geom_vline(aes(xintercept = breaks.cluster-3, color = "cluster"))
```
